---
title: "Reproducible assignment 1 - activity monitoring"
author: "fruzsi"
date: "19 November 2016"
output: 
    html_document:
        keep_md: true
---

Loading packages

```{r lib}
library(dplyr)
library(ggplot2)
```


# Setting parameters, loading data, and having a peak

```{r dataload}
DIR <-'C:\\Users\\fishie\\Desktop\\coursera\\repoClone\\datasciencecoursera\\replic_assignmnent1'
fname <- 'activity.csv'
data <- read.csv(paste(DIR,'\\', fname, sep=''))
head(data)
summary(data)
```

# Mean total number of steps taken per day

Transform date column from character to Date object 

```{r}
data$date <- as.Date(data$date)
```

Calculate number of steps per day

```{r stepsum}
steps_stat <- data %>%
    group_by(date) %>%
    filter(!is.na(steps)) %>%
    summarize(stepcount = sum(steps))

steps_stat
```

Plot step counts per day

```{r stepsumplot}
p <- ggplot(steps_stat, aes(stepcount))
p + geom_histogram(binwidth = 800, fill='darkblue') +
    labs(title='total steps per days', x='steps per day', y='frequency')
```

Mean and median of steps taken a day

```{r}
# Mean:
mean(steps_stat$stepcount)
# Median:
median(steps_stat$stepcount)
```

# Discover average daily pattern

Calculate average steps per intervals

```{r stepsavg}
steps_stat <- data %>%
    group_by(interval) %>%
    filter(!is.na(steps)) %>%
    summarize(stepavg = mean(steps))

steps_stat
```

Plot average daily steps

```{r stepsavgplot}
p <- ggplot(steps_stat, aes(x=interval, y=stepavg))
p + geom_line(col='darkblue', size=1) +
    labs(title='average daily steps', x='interval', y='average steps per day')
```

Find the 5-minute interwal that on average contains the most number of steps

```{r}
steps_stat[which.max(steps_stat$stepavg),]
```


# Imputing missing values

number of missing values in the dataset

```{r}
sum(is.na(data$steps))
```

```{r imput}
na.ind <- is.na(data$steps)
data.means <- tapply(data$steps, data$interval, mean, na.rm=T, simplify=T)
data.imputed <- data
data.imputed$steps[na.ind] <- data.means[as.character(data$interval[na.ind])]
sum(is.na(data.imputed$steps))

```

Histogram of total number of steps per days - imputed data

```{r}
steps_stat <- data.imputed %>%
    group_by(date) %>%
    filter(!is.na(steps)) %>%
    summarize(stepcount = sum(steps))

# plot
p <- ggplot(steps_stat, aes(stepcount))
p + geom_histogram(binwidth = 800, fill='darkred') +
    labs(title='total steps per days - imputed data', x='steps per day', y='frequency')
```

Mean and median of steps taken a day - imputed data

```{r}
# Mean - imputed data:
mean(steps_stat$stepcount)
# Median - imputed data:
median(steps_stat$stepcount)
```
-> imputing values didn't make very much of a difference; moved the median closer (to the same) to the mean, but that's not unexpected, after imputing mean values

# Differences between activity patterns on weekdays vs weekends

Create factor variable for weekday and weekend

```{r}
data.imputed <- data.imputed %>%
    mutate(weekpart = ifelse(weekdays(date)=='Saturday' | 
                                  weekdays(date)=='Sunday',
                              'weekend', 'weekday'))

# statistics
steps_stat <- data.imputed %>%
    group_by(interval,weekpart) %>%
    summarize(stepavg = mean(steps))

head(steps_stat)
# plot
p <- ggplot(steps_stat, aes(x=interval, y=stepavg, color=weekpart))
p + geom_line(size=1) +
    facet_wrap(~weekpart,nrow=2,ncol=1) +
    labs(title='average daily steps, by weekpart', x='interval', 
         y='average steps per day')
```


-> Subject seems to be more active during workdays in the morning, but then a bit more active during the day in the weekends

